---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setopt, echo = FALSE, results = 'hide'}
opts_chunk$set(echo = TRUE, fig.height = 4)
Sys.setlocale("LC_ALL", 'en_US.UTF-8')
```

# "Reproducible Research: Peer Assessment 1"

#### author: Micael Eiji Nagai

## Loading and preprocessing the data
Loading/installing some packages and acquiring and processing the data:

```{r preprocessing}
file <- 'http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'
if(!file.exists('.//activity.zip')) download.file( url = file, destfile = ".//activity.zip")
if(!file.exists( './/data//activity.csv')) unzip('.//activity.zip', exdir = './/data')
if(!require('data.table'))install.packages('data.table');library('data.table')
if(!require('ggplot2'))install.packages('ggplot2');library('ggplot2')
dtable <- read.table( './/data//activity.csv', sep = ',', header = TRUE)
dtable$date <- as.Date( dtable$date, format = "%Y-%m-%d")
dtable <- data.table(dtable)
```


## What is mean total number of steps taken per day?

The total number of steps taken per day:
```{r}
dtable[, list(sum.steps=sum(steps, na.rm = TRUE)), by = date]
```

A histogram showing these information:
```{r}
gg <- ggplot(dtable[, list(sum.steps=sum(steps, na.rm = TRUE)), by = date])
gg <- gg+geom_histogram(aes(x = sum.steps),binwidth = 400)+theme_classic()
gg <- gg+labs(x='Steps')
print(gg)
```

Some summary, including mean and median of the steps:
```{r}
dtable[, list(sum.steps=sum(steps, na.rm = TRUE)), by = date][,summary(sum.steps)]
```

## What is the average daily activity pattern?

Daily activity pattern, through 5 min interval, measured as a average from the days. 

```{r}
gg <- ggplot( data = dtable[, list( av = mean(steps,na.rm = TRUE)), by = interval], aes(x=interval, y=av))
gg <- gg + geom_line( ) + labs(y='Mean steps',x='Daily interval (5 min)')
gg <- gg + theme_classic()
print(gg)
```

On average the interval that the individual made more steps was:

```{r}
dtable[, list( av = mean(steps,na.rm = TRUE)), by = interval][which(av==max(av))]
```

## Imputing missing values

The amount of NA's on steps variable of the data:

```{r}
dtable[is.na(steps), length(steps)]
```

To fill the NA's value I will use the median for the 5 minutes interval.
The new dataset will be named nafilled.

```{r}
nafilled <- copy(dtable)
setkey(nafilled,interval)

steps.na <- which(is.na(nafilled[,steps]))
interval.na <- nafilled[steps.na,interval]
data.tmp <- dtable[,list(av=median(steps,na.rm=TRUE)),by=interval]
nafilled[steps.na, `:=`  (steps = data.tmp[interval==interval.na,av])]
rm(data.tmp)
```

And the histogram for the filled NA data

```{r}
gg <- ggplot( data = nafilled, aes( x = steps))
gg <- gg+geom_histogram(binwidth = 30)
gg <- gg+theme_classic()
print(gg)
```

Changing the NA's values for the median of steps of the 5 minutes interval that it belong inserted a lot of 0's values, so the counts of it increased a lot.

## Are there differences in activity patterns between weekdays and weekends?

Creating a factor variable with the value of weekday and weekend.
And then creating a temporary data.table with the average steps by the 5 minutes interval for the weekday and weekend.

```{r}
weekend <- c('saturday','sunday')
nafilled[, week:= as.factor(ifelse(tolower(weekdays(date)) %in% weekend, 'weekend','weekday'))]
data.tmp <- nafilled[, list(av= mean(steps)),by = list(interval,week)]
```
Plot the average steps on the 5 minutes interval for the weekday and weekend.

```{r, fig.height = 6}
gg <- ggplot(data=data.tmp,aes( x = interval, y = av))
gg <- gg+geom_line() + facet_grid(week~.)
gg <- gg+theme_bw() + labs( x='5 minutes interval', y='Steps')
print(gg)
```
